---
title: "Analysis of school drug zone arrests in Connecticut"
author: "Andrew Ba Tran"
date: "June 5, 2015"
output: html_document
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
require(lubridate)
require(stringi)
require(gridExtra)
require(gtools)
require(ggplot2)
require(rgdal)
require(scales)
require(ggmap)
require(dplyr)
require(Cairo)
require(gpclib)
require(maptools)
require(reshape)
require(knitr)

gpclibPermit()
gpclibPermitStatus()
```

#### Before we begin: our data has `r I(nrow(arrests))` rows and `r I(ncol(arrests))` columns.

#### How many drug arrests have there been for 21a-267 (c),  21a-278 (b), 21a-279 (d) since 1999?

```{r results='asis', out.width='910px', fig.width=10, fig.height=5, warning=FALSE, message=FALSE, dpi=50}
arrests <- read.csv("arrests.csv")

# Clean up the dates
arrests$Date <- dmy(arrests$ARREST.DATE)
arrests$Year <- year(arrests$Date)

# Plot out year
plot(table(arrests$Year))
```

#### What's the racial breakdown for thoses arrests?

```{r results='asis', out.width='910px', fig.width=10, fig.height=5, warning=FALSE, message=FALSE, dpi=50}

# Racial breakdown of arrests by year
race_year <- data.frame(table(arrests$Year,arrests$RACE))
colnames(race_year) <- c("Year", "Race", "Arrests")
ggplot(race_year, aes(Year, Arrests, group=Race, colour=Race)) +
  geom_path(alpha=0.5) +
  ggtitle("Race of those arrested since 1999") +
  theme_minimal()

# And by percent
ggplot(race_year, aes(Year, Arrests, group=Race, fill=Race)) + geom_area(position="fill") +
  ggtitle("Race of those arrested since 1999 by percent")
```

####  Which towns have the most arrests?

```{r results='asis'}
town_arrests <- data.frame(table(arrests$Town))
colnames(town_arrests) <- c("id", "Total.Arrests")
town_arrests <- town_arrests[order(-town_arrests$Total.Arrests),]
kable(head(town_arrests, 10))

```

#### Let's adjust for town population

```{r results='asis', out.width='910px', fig.width=10, fig.height=5, warning=FALSE, message=FALSE, dpi=50}

# Bring in town population
townpop <- read.csv("townsmap/townpopulation.csv")
colnames(townpop) <- c("id","Population")
townpop$id <- toupper(townpop$id)

# Have to make adjustments to neighborhood names to match town names

town_arrests$id <- gsub("BANTAM", "LITCHFIELD", town_arrests$id)
town_arrests$id <- gsub("BELLE HAVEN", "GREENWICH", town_arrests$id)
town_arrests$id <- gsub("BRANCHVILLE", "RIDGEFIELD", town_arrests$id)
town_arrests$id <- gsub("CANNONDALE", "WILTON", town_arrests$id)
town_arrests$id <- gsub("CENTERVILLE", "HAMDEN", town_arrests$id)
town_arrests$id <- gsub("CHESHIRE HEIGHTS", "CHESHIRE", town_arrests$id)
town_arrests$id <- gsub("DANIELSON", "WINDHAM", town_arrests$id)
town_arrests$id <- gsub("DOBSONVILLE", "VERNON", town_arrests$id)
town_arrests$id <- gsub("EAST SUFFIELD", "SUFFIELD", town_arrests$id)
town_arrests$id <- gsub("FITCHVILLE", "BOZRAH", town_arrests$id)
town_arrests$id <- gsub("GOSHEN CENTER", "GOSHEN", town_arrests$id)
town_arrests$id <- gsub("GROTON CITY OF", "GROTON", town_arrests$id)
town_arrests$id <- gsub("JEWETT CITY", "GRISWOLD", town_arrests$id)
town_arrests$id <- gsub("MOOSUP", "PLAINFIELD", town_arrests$id)
town_arrests$id <- gsub("NORTH COLEBROOK", "COLEBROOK", town_arrests$id)
town_arrests$id <- gsub("OAKVILLE", "WATERTOWN", town_arrests$id)
town_arrests$id <- gsub("POQUONOCK", "WINDSOR", town_arrests$id)
town_arrests$id <- gsub("PUTNAM HEIGHTS", "PUTNAM", town_arrests$id)
town_arrests$id <- gsub("ROCKVILLE", "VERNON", town_arrests$id)
town_arrests$id <- gsub("SOUTH CANAAN", "CANAAN", town_arrests$id)
town_arrests$id <- gsub("STORRS", "MANSFIELD", town_arrests$id)
town_arrests$id <- gsub("UNCASVILLE", "MONTVILLE", town_arrests$id)
town_arrests$id <- gsub("WAREHOUSE POINT", "WINDSORVILLE", town_arrests$id)
town_arrests$id <- gsub("WEST MORRIS", "LITCHFIELD", town_arrests$id)

town_arrests <- data.frame(tapply(town_arrests$Total.Arrests, town_arrests$id, sum))
town_arrests$id <- row.names(town_arrests)
colnames(town_arrests) <- c("Total.Arrests", "id")

towns_arrests_pop <- left_join(townpop, town_arrests)
towns_arrests_pop$Total.Arrests[is.na(towns_arrests_pop$Total.Arrests)] <-0
towns_arrests_pop$id <- stri_trans_general(towns_arrests_pop$id, id="Title")
towns_arrests_pop$Per10kResidents <- (towns_arrests_pop$Total.Arrests/towns_arrests_pop$Population)*10000
towns_arrests_pop$Per10kResidents <- round(towns_arrests_pop$Per10kResidents, digits=2)
towns_arrests_pop <- towns_arrests_pop[order(-towns_arrests_pop$Per10kResidents),]

kable(head(towns_arrests_pop, 10))

```

#### Ansonia? Really? Ok let's map it out

```{r results='asis', out.width='910px', fig.width=10, fig.height=5, warning=FALSE, message=FALSE, dpi=50}
towntracts <- readOGR(dsn="townsmap", layer="towns")
towntracts <- fortify(towntracts, region="NAME10")

town_arr_Data <- left_join(towntracts, towns_arrests_pop)

p2 <- ggplot() +
  geom_polygon(data = town_arr_Data, aes(x=long, y=lat, group=group, 
                                    fill=Per10kResidents), color = "black", size=0.2) +
  coord_map() +
  scale_fill_distiller(type="seq", palette = "Greens", breaks=pretty_breaks(n=5)) +
  theme_nothing(legend=TRUE) +
  labs(title="Arrests between 1999 and 2014 per 10,000 residents", fill="")
p2

```

#### Break out the arrests by charge

```{r results='asis', out.width='910px', fig.width=10, fig.height=5, warning=FALSE, message=FALSE, dpi=50}
charge_year <- data.frame(table(arrests$Year,arrests$ORIGINAL.STATUTE))
colnames(charge_year) <- c("Year", "Charge", "Arrests")
ggplot(charge_year, aes(Year, Arrests, group=Charge, colour=Charge)) +
  geom_path(alpha=0.5) +
  ggtitle("Types of charges since 1999") +
  theme_minimal()

# Percent of arrests by year
ggplot(charge_year, aes(Year, Arrests, group=Charge, fill=Charge)) + geom_area(position="fill")

# OK, all of the above but specifically at D

just_d <- subset(arrests, ORIGINAL.STATUTE=="21a-279(d)")
plot(table(just_d$Year))

race_year_d <- data.frame(table(just_d$Year,just_d$RACE))
colnames(race_year_d) <- c("Year", "Race", "Arrests")
ggplot(race_year_d, aes(Year, Arrests, group=Race, colour=Race)) +
  geom_path(alpha=0.5) +
  ggtitle("Race of those arrested for Possession since 1999") +
  theme_minimal()

# And by percent
ggplot(race_year_d, aes(Year, Arrests, group=Race, fill=Race)) + geom_area(position="fill")
```

# Which towns have the most arrests for Possession?


```{r results='asis'}

# Which towns have the most arrests?
town_arrests_d <- data.frame(table(just_d$Town))
colnames(town_arrests_d) <- c("id", "Total.Arrests")
town_arrests_d <- town_arrests_d[order(-town_arrests_d$Total.Arrests),]
kable(head(town_arrests_d, 10))

```

# Which towns have the most arrests for Possession (after adjusting for population)?


```{r results='asis'}
town_arrests_d$id <- gsub("BANTAM", "LITCHFIELD", town_arrests_d$id)
town_arrests_d$id <- gsub("BELLE HAVEN", "GREENWICH", town_arrests_d$id)
town_arrests_d$id <- gsub("BRANCHVILLE", "RIDGEFIELD", town_arrests_d$id)
town_arrests_d$id <- gsub("CANNONDALE", "WILTON", town_arrests_d$id)
town_arrests_d$id <- gsub("CENTERVILLE", "HAMDEN", town_arrests_d$id)
town_arrests_d$id <- gsub("CHESHIRE HEIGHTS", "CHESHIRE", town_arrests_d$id)
town_arrests_d$id <- gsub("DANIELSON", "WINDHAM", town_arrests_d$id)
town_arrests_d$id <- gsub("DOBSONVILLE", "VERNON", town_arrests_d$id)
town_arrests_d$id <- gsub("EAST SUFFIELD", "SUFFIELD", town_arrests_d$id)
town_arrests_d$id <- gsub("FITCHVILLE", "BOZRAH", town_arrests_d$id)
town_arrests_d$id <- gsub("GOSHEN CENTER", "GOSHEN", town_arrests_d$id)
town_arrests_d$id <- gsub("GROTON CITY OF", "GROTON", town_arrests_d$id)
town_arrests_d$id <- gsub("JEWETT CITY", "GRISWOLD", town_arrests_d$id)
town_arrests_d$id <- gsub("MOOSUP", "PLAINFIELD", town_arrests_d$id)
town_arrests_d$id <- gsub("NORTH COLEBROOK", "COLEBROOK", town_arrests_d$id)
town_arrests_d$id <- gsub("OAKVILLE", "WATERTOWN", town_arrests_d$id)
town_arrests_d$id <- gsub("POQUONOCK", "WINDSOR", town_arrests_d$id)
town_arrests_d$id <- gsub("PUTNAM HEIGHTS", "PUTNAM", town_arrests_d$id)
town_arrests_d$id <- gsub("ROCKVILLE", "VERNON", town_arrests_d$id)
town_arrests_d$id <- gsub("SOUTH CANAAN", "CANAAN", town_arrests_d$id)
town_arrests_d$id <- gsub("STORRS", "MANSFIELD", town_arrests_d$id)
town_arrests_d$id <- gsub("UNCASVILLE", "MONTVILLE", town_arrests_d$id)
town_arrests_d$id <- gsub("WAREHOUSE POINT", "WINDSORVILLE", town_arrests_d$id)
town_arrests_d$id <- gsub("WEST MORRIS", "LITCHFIELD", town_arrests_d$id)

town_arrests_d <- data.frame(tapply(town_arrests_d$Total.Arrests, town_arrests_d$id, sum))

town_arrests_d$id <- row.names(town_arrests_d)
colnames(town_arrests_d) <- c("Total.Arrests", "id")

towns_arrests_pop_d <- left_join(townpop, town_arrests_d)
towns_arrests_pop_d$Total.Arrests[is.na(towns_arrests_pop_d$Total.Arrests)] <-0
towns_arrests_pop_d$id <- stri_trans_general(towns_arrests_pop_d$id, id="Title")
towns_arrests_pop_d$Per10kResidents <- (towns_arrests_pop_d$Total.Arrests/towns_arrests_pop_d$Population)*10000
towns_arrests_pop_d$Per10kResidents <- round(towns_arrests_pop_d$Per10kResidents, digits=2)
towns_arrests_pop_d <- towns_arrests_pop_d[order(-towns_arrests_pop_d$Per10kResidents),]

kable(head(towns_arrests_pop_d, 10))

town_arr_Data_d <- left_join(towntracts, towns_arrests_pop_d)

p3 <- ggplot() +
  geom_polygon(data = town_arr_Data_d, aes(x=long, y=lat, group=group, 
                                         fill=Per10kResidents), color = "black", size=0.2) +
  coord_map() +
  scale_fill_distiller(type="seq", palette = "Greens", breaks=pretty_breaks(n=5)) +
  theme_nothing(legend=TRUE) +
  labs(title="Possession arrests between 1999 and 2014 per 10,000 residents", fill="")
p3
```

### What’s the racial breakdown in towns with the most Possession convictions?

```{r results='asis', out.width='910px', fig.width=10, fig.height=10, warning=FALSE, message=FALSE, dpi=50}
# New Haven
arrests_d_New.Haven <- subset(arrests, Town=="NEW HAVEN")
arrests_d_New.Haven.year <- data.frame(table(arrests_d_New.Haven$Year, arrests_d_New.Haven$RACE))
colnames(arrests_d_New.Haven.year) <- c("Year", "Race", "Arrests")

nh1a <- ggplot(arrests_d_New.Haven.year, aes(Year, Arrests, group=Race, colour=Race)) +
  geom_path(alpha=0.5) +
ggtitle("Total") +
  theme(legend.position="top")

nh2a <- ggplot(arrests_d_New.Haven.year, aes(Year, Arrests, group=Race, fill=Race)) + geom_area(position="fill") + 
  ggtitle("Percent")  +
  theme(legend.position="top")

grid.arrange(nh1a, nh2a, ncol=1, main="Race of those arrested for Possession in New Haven")

# Hartford
arrests_d_Hartford <- subset(arrests, Town=="HARTFORD")
arrests_d_Hartford.year <- data.frame(table(arrests_d_Hartford$Year, arrests_d_Hartford$RACE))
colnames(arrests_d_Hartford.year) <- c("Year", "Race", "Arrests")

h1a <- ggplot(arrests_d_Hartford.year, aes(Year, Arrests, group=Race, colour=Race)) +
  geom_path(alpha=0.5) +
ggtitle("Total") +
  theme(legend.position="top")

h2a <- ggplot(arrests_d_Hartford.year, aes(Year, Arrests, group=Race, fill=Race)) + geom_area(position="fill") + 
  ggtitle("Percent")  +
  theme(legend.position="top")

grid.arrange(h1a, nh2a, ncol=1, main="Race of those arrested for Possession in Hartford")

# Waterbury
arrests_d_Waterbury <- subset(arrests, Town=="WATERBURY")
arrests_d_Waterbury.year <- data.frame(table(arrests_d_Waterbury$Year, arrests_d_Waterbury$RACE))
colnames(arrests_d_Waterbury.year) <- c("Year", "Race", "Arrests")

h1a <- ggplot(arrests_d_Waterbury.year, aes(Year, Arrests, group=Race, colour=Race)) +
  geom_path(alpha=0.5) +
ggtitle("Total") +
  theme(legend.position="top")

h2a <- ggplot(arrests_d_Waterbury.year, aes(Year, Arrests, group=Race, fill=Race)) + geom_area(position="fill") + 
  ggtitle("Percent")  +
  theme(legend.position="top")

grid.arrange(h1a, nh2a, ncol=1, main="Race of those arrested for Possession in Waterbury")

# Ansonia
arrests_d_Ansonia <- subset(arrests, Town=="ANSONIA")
arrests_d_Ansonia.year <- data.frame(table(arrests_d_Ansonia$Year, arrests_d_Ansonia$RACE))
colnames(arrests_d_Ansonia.year) <- c("Year", "Race", "Arrests")

h1a <- ggplot(arrests_d_Ansonia.year, aes(Year, Arrests, group=Race, colour=Race)) +
  geom_path(alpha=0.5) +
ggtitle("Total") +
  theme(legend.position="top")

h2a <- ggplot(arrests_d_Ansonia.year, aes(Year, Arrests, group=Race, fill=Race)) + geom_area(position="fill") + 
  ggtitle("Percent")  +
  theme(legend.position="top")

grid.arrange(h1a, nh2a, ncol=1, main="Race of those arrested for Possession in Ansonia")

# Norwalk
arrests_d_Norwalk <- subset(arrests, Town=="NORWALK")
arrests_d_Norwalk.year <- data.frame(table(arrests_d_Norwalk$Year, arrests_d_Norwalk$RACE))
colnames(arrests_d_Norwalk.year) <- c("Year", "Race", "Arrests")

h1a <- ggplot(arrests_d_Norwalk.year, aes(Year, Arrests, group=Race, colour=Race)) +
  geom_path(alpha=0.5) +
ggtitle("Total") +
  theme(legend.position="top")

h2a <- ggplot(arrests_d_Norwalk.year, aes(Year, Arrests, group=Race, fill=Race)) + geom_area(position="fill") + 
  ggtitle("Percent")  +
  theme(legend.position="top")

grid.arrange(h1a, nh2a, ncol=1, main="Race of those arrested for Possession in Norwalk")

arrests_urban <- subset(arrests, (Town=="BRIDGEPORT" | Town=="HARTFORD" | Town=="NEW HAVEN" |
                               Town=="NEW BRITAIN" | Town=="WEST HAVEN" | Town=="NEW LONDON" |
                               Town=="WATERBURY" | Town=="NORWALK" | Town=="WATERBURY" |
                               Town=="NORWALK" | Town=="ANSONIA" | Town=="STAMFORD"))

arrests_urban.year <- data.frame(table(arrests_urban$Year, arrests_urban$RACE))
colnames(arrests_urban.year) <- c("Year", "Race", "Arrests")

adu1 <- ggplot(arrests_urban.year, aes(Year, Arrests, group=Race, colour=Race)) +
  geom_path(alpha=0.5) +
ggtitle("Total") +
  theme(legend.position="top")

adu2 <- ggplot(arrests_urban.year, aes(Year, Arrests, group=Race, fill=Race)) + geom_area(position="fill") + 
  ggtitle("Percent")  +
  theme(legend.position="top")

grid.arrange(adu1, adu2, ncol=1, main="Race of those arrested in Urban towns")



arrests_suburban <- subset(arrests, !(Town=="BRIDGEPORT" | Town=="HARTFORD" | Town=="NEW HAVEN" |
                               Town=="NEW BRITAIN" | Town=="WEST HAVEN" | Town=="NEW LONDON" |
                               Town=="WATERBURY" | Town=="NORWALK" | Town=="WATERBURY" |
                               Town=="NORWALK" | Town=="ANSONIA" | Town=="STAMFORD"))

arrests_suburban.year <- data.frame(table(arrests_suburban$Year, arrests_suburban$RACE))
colnames(arrests_suburban.year) <- c("Year", "Race", "Arrests")

ads1 <- ggplot(arrests_suburban.year, aes(Year, Arrests, group=Race, colour=Race)) +
  geom_path(alpha=0.5) +
ggtitle("Total") +
  theme(legend.position="top")

ads2 <- ggplot(arrests_suburban.year, aes(Year, Arrests, group=Race, fill=Race)) + geom_area(position="fill") + 
  ggtitle("Percent")  +
  theme(legend.position="top")

grid.arrange(ads1, ads2, ncol=1, main="Race of those arrested in Suburban towns")


arrests_d_urban <- subset(just_d, (Town=="BRIDGEPORT" | Town=="HARTFORD" | Town=="NEW HAVEN" |
                               Town=="NEW BRITAIN" | Town=="WEST HAVEN" | Town=="NEW LONDON" |
                               Town=="WATERBURY" | Town=="NORWALK" | Town=="WATERBURY" |
                               Town=="NORWALK" | Town=="ANSONIA" | Town=="STAMFORD"))

arrests_d_urban.year <- data.frame(table(arrests_d_urban$Year, arrests_d_urban$RACE))
colnames(arrests_d_urban.year) <- c("Year", "Race", "Arrests")

ady1 <- ggplot(arrests_d_urban.year, aes(Year, Arrests, group=Race, colour=Race)) +
  geom_path(alpha=0.5) +
ggtitle("Total") +
  theme(legend.position="top")

ady2 <- ggplot(arrests_d_urban.year, aes(Year, Arrests, group=Race, fill=Race)) + geom_area(position="fill") + 
  ggtitle("Percent")  +
  theme(legend.position="top")

grid.arrange(ady1, ady2, ncol=1, main="Race of those arrested for Possession in Urban towns")

arrests_d_suburban <- subset(just_d, !(Town=="BRIDGEPORT" | Town=="HARTFORD" | Town=="NEW HAVEN" |
                               Town=="NEW BRITAIN" | Town=="WEST HAVEN" | Town=="NEW LONDON" |
                               Town=="WATERBURY" | Town=="NORWALK" | Town=="WATERBURY" |
                               Town=="NORWALK" | Town=="ANSONIA" | Town=="STAMFORD"))

arrests_d_suburban.year <- data.frame(table(arrests_d_suburban$Year, arrests_d_suburban$RACE))
colnames(arrests_d_suburban.year) <- c("Year", "Race", "Arrests")

ady1 <- ggplot(arrests_d_suburban.year, aes(Year, Arrests, group=Race, colour=Race)) +
  geom_path(alpha=0.5) +
ggtitle("Total") +
  theme(legend.position="top")

ady2 <- ggplot(arrests_d_suburban.year, aes(Year, Arrests, group=Race, fill=Race)) + geom_area(position="fill") + 
  ggtitle("Percent")  +
  theme(legend.position="top")

grid.arrange(ady1, ady2, ncol=1, main="Race of those arrested for Possession in Suburban towns")
```
